---
title: "Summary statitistics for the Orange data set"
author: "Malte Berneaud-Koetz"
date: "22 February 2016"
output: pdf_document
---

## Tree growth over time 

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(ggplot2)
source("../summary_statistics.R")
```

```{r}
ggplot(Orange, aes(x = age, y = circumference, colour = Tree)) + geom_line()
```

## Measures of central tendency: means

Overall, the mean circumference of trees across the entire study period and across all trees is 115.86 centimeters. The means of the individual trees is found in the table below:

| Tree   | Mean of circumference |
|--------|-----------------------|
| Tree 1 | 99.6 cm               |
| Tree 2 | 135.29 cm             |
| Tree 3 | 94.0 cm               |
| Tree 4 | 139.29 cm             |
| Tree 5 | 111.14 cm             |


## Measures of central tendency: medians

I calculated the medians across all trees, which is 115 cm. Additionally, I calculated the medians for each of the individual trees, as given in the table below:

| Tree   | Median of circumference |
|--------|-------------------------|
| Tree 1 | 115 cm                  |
| Tree 2 | 156 cm                  |
| Tree 3 | 108 cm                  |
| Tree 4 | 167 cm                  |
| Tree 5 | 156 cm                  |

## Measures of dispersion

I calculated the standard deviation for the entire data set, which is 57.49. Moreover, I presented the standard deviations among the different trees below:

| Tree   | Std. dev. of circumference |
|--------|----------------------------|
| Tree 1 | 43.29 cm                   |
| Tree 2 | 66.24 cm                   |
| Tree 3 | 42.98 cm                   |
| Tree 4 | 71.90 cm                   |
| Tree 5 | 58.86 cm                   |

## All summary statistics in one graph

```{r}
barplot(summary.matrix, beside = TRUE,
        main = "Summary statistics for circumference of all trees",
        xlab = "Name of summary statistic",
        ylab = "Tree trunk circumference",
        names.arg = names(summary.df$`Tree ID`))
```

## Questions & Answers
Questions? 

Points of critique? 

Suggestions? 
