---
title: "Descriptive statistics of Life Cycle Savings R Database"
author: "Jonas Markgraf"
date: "22 February 2016"
output: pdf_document
---
This file contains descriptive analysis - measures of central tendency as well as measures of dispersion - of the LifeCycleSavings R dataset. Furthermore, it will be looked into correlations between different variables. A particular focus will be on the two variables *Aggregated Savings (sr)* and *Per-Capita Income (dpi)*, thus testing the hypothesis that aggregated savings increase with higher per-capita income.

# Description of dataset and variables of interest

The LifeCycleSavings dataset is a data frame containing information on the savings ratio between 1960 and 1970 over 50 countries. It includes the following 5 variables (taken from R):

- _**sr**_:     aggregate personal savings

- _**pop15**_:  % of population under 15

- _**pop75**_:  %  of population over 75

- _**dpi**_:    real per-capital disposable income

- _**ddpi**_:   % growth rate of dpi.

# Guide to R code

The R code in this folder conducts a number of descriptive statistics, namely measures of central tendency and measures of dispersion. Furthermore, correlations between variables were analyzed as a first step to identify potential explanatory variables (per-capita GDP; per-capita GDP growth; demographic factors) for the aggregated personal savings.

Measures of central tendency: histograms, mean, median

Measures of dispersion: standard deviation, range, interquartile range, boxplots

Correlations: plots; significance tests

# Descriptive Statistics

```{r, include=FALSE}
# Load packages
library(ggplot2)
library(knitr)
library(dplyr)
library(magrittr)

# Load dataset
library(datasets)
data(LifeCycleSavings)

# set working directory
setwd("/Users/jonasmarkgraf/Desktop/DataScienceAssignment1/JonasFiles")

# Detailed information about dataset, incl. description of variables
?LifeCycleSavings
```

## Measures of central tendency

First, we have a look at the key measures of central tendency starting with the means of the five variables:

```{r}
for (i in 1:5) {
  LifeCycleSavings[, i] %>%
    mean() %>%
    paste(names(LifeCycleSavings)[i], ., "\n") %>%
    cat()
}
```
As a next step, the medians are calculated:
```{r}
for (i in 1:5) {
  LifeCycleSavings[, i] %>%
    median() %>%
    paste(names(LifeCycleSavings)[i], ., "\n") %>%
    cat()
}
```
After the calculating mean and median, we have a look at the histograms of 2 key variables - Per-Capita Income and Aggregated Savings - in order to get an idea of the general distribution of those two variables:

```{r}
par(mfcol = c(1, 2))
hist(LifeCycleSavings$dpi, 
     main="Per-Capita Income Distribution", 
     xlab="Per-Capita Income", 
     ylab="Number of Countries")
hist(LifeCycleSavings$sr, 
     main="Aggregated Savings Distribution", 
     xlab="Aggregated Savings", 
     ylab="Number of Countries")
```

## Measures of dispersion

In a next step, we analyse the measures of dispersion for the *LifeCycleSavings* R Dataset. We namely calculate the standard deviation for all five variables.

```{r}
## loop for standard deviation
for (i in 1:5) {
  LifeCycleSavings[, i] %>%
    sd() %>%
    paste(names(LifeCycleSavings)[i], ., "\n") %>%
    cat()
}
```

Furthermore, we look at the range and the interquartile range of the the key variables *aggregated savings* and *per-capita income*.

```{r}
## range for 2 key variables
range(LifeCycleSavings$sr)
range(LifeCycleSavings$dpi)

## interquartile range for 2 key variables
IQR(LifeCycleSavings$sr)
IQR(LifeCycleSavings$dpi)
```
Eventually, we graphically plot the measures of central tendency and dispersion for our two key variables with two boxplots, which provides us with a graphical presentation of mean, range and interquartile range.
```{r}
par(mfcol = c(1, 2))
boxplot(LifeCycleSavings$sr, main = "Aggregated Savings", ylab = "Savings")
boxplot(LifeCycleSavings$dpi, main = "Per-Capita Income", ylab = "Income in $")
```

# Joint distributions

In order to identify potential explanatory variables for the aggregated savings, we look into correlations between the the demographic variables (*pop15* and *pop75*) as well as different specifications of the economic situation in the respective country (absolute (*dpi*) and relative (*ddpi*) measures of per-capita income).

## Demographic factors

First we start with the two demographic variables: The share of population older than 75 years and younger than 15 years.
```{r, "Demographic Factors"}
ggplot(LifeCycleSavings, aes(pop15, sr)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Correlation between Share of Population under 15 Years and 
          Aggregated Savings") +
  xlab("Share of Population under 15 Years (%)") +
  ylab("Aggregated Savings")

ggplot(LifeCycleSavings, aes(pop75, sr)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Correlation between Share of Population older than 75 Years and
          Aggregated Savings") +
  xlab("Share of Population over 75 Years (%)") +
  ylab("Aggregated Savings")
```

## Income factors

Secondly, we look at the correlation of different income specifications (absolute per-capita income level *dpi* and relative per-capita income growth *ddpi*).
```{r}
ggplot(LifeCycleSavings, aes(dpi, sr)) + 
  geom_point() + 
  geom_smooth() + 
  ggtitle("Savings-Per Capita GDP Correlation") +
  xlab("Per Capita Income") +
  ylab("Aggregated savings")

ggplot(LifeCycleSavings, aes(ddpi, sr)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Correlation between Per-Capita Income Growth and Aggregated Savings") +
  xlab("Per-Capita Income Growth (%)") +
  ylab("Aggregated Savings")
```